---
title: "LDA"
output: html_document
date: '2022-11-26'
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      error = FALSE, 
                      message = FALSE)
```

```{r, include = FALSE}
pacman::p_load(tidyverse, readr, stringr, quanteda, quanteda.textstats, rvest, 
               tibble, xml2, manifestoR, topicmodels, tidytext, stm, kableExtra,
               janitor)

path_repo <- "/Users/janinedevera/Documents/School/MDS 2021-2023/Semester 3/2 Text as Data/Project/antitRust"
path_charts = paste0(path_repo, "/plots")
```

```{r, include = FALSE}
charts.theme <- theme(axis.title.y.left = element_text(size = 12, margin = margin(r = 15)),
                      axis.title.y.right = element_text(size = 12, margin = margin(l = 15)),
                      axis.title.x = element_text(size = 12, margin = margin(t = 15, b = 8)),
                      axis.text.x = element_text(size = 12),
                      axis.text.y = element_text(size = 12),
                      axis.ticks = element_blank(),
                      axis.line.x = element_line("transparent", size = 0.5), 
                      axis.line.y = element_line("transparent", size = 0.5),
                      panel.border = element_rect(color = "#a3a3a3", fill = "transparent"),
                      panel.background = element_rect(fill = "white", color = "white"),
                      panel.grid.major = element_line(color = "#d4d4d4", linetype = 2),
                      plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
                      plot.subtitle = element_text(size = 10, face = "italic", hjust = 0.5, margin = margin(b = 15)),
                      plot.caption = element_text(size = 10, hjust = 0, face = "italic"),
                      strip.background = element_rect(fill = "transparent"),
                      strip.text = element_text(size = 12),
                      legend.key=element_blank())
```

### I. Descriptives
```{r}
raw <- read_csv(paste0(path_repo, "/data/df_clean_eng.csv"), show_col_types = FALSE)[,-1] %>% 
 tibble::rownames_to_column("doc_id")
```

Platforms:
```{r}
raw %>% 
 group_by(company_search) %>%
 summarise(counts = n()) %>% 
 arrange(desc(counts)) 
 #tabyl(company) %>% 
 #arrange(desc(n)) %>% 
 #adorn_totals("row") %>%
 #adorn_pct_formatting(digits = 0) %>% 
 #kable(col.names = c("Company", "# Articles", "Percent")) %>% 
 #kable_styling(bootstrap_options = c("striped", "condensed", "hover"), 
 #               full_width = F) %>% 
 #row_spec(6, bold = T)

```

Years:
```{r}
raw %>% 
 group_by(year) %>%
 summarise(counts = n()) %>% 
 arrange(desc(year))
 #tabyl(year) %>% 
 #arrange(desc(year)) %>% 
 #adorn_totals("row") %>%
 #adorn_pct_formatting(digits = 0) %>% 
 #kable(col.names = c("Year", "# Articles", "Percent")) %>% 
 #kable_styling(bootstrap_options = c("striped", "condensed", "hover"), 
 #               full_width = F) %>% 
 #row_spec(15, bold = T)
```

Countries:
```{r}
raw %>% 
 filter(country != "None") %>% 
 group_by(country) %>%
 summarise(counts = n()) %>%
 arrange(desc(counts))
 #tabyl(country) %>% 
 #arrange(desc(n)) %>% 
 #adorn_totals("row") %>%
 #adorn_pct_formatting(digits = 0) %>% 
 #kable(col.names = c("Country", "# Articles", "Percent")) %>% 
 #kable_styling(bootstrap_options = c("striped", "condensed", "hover"), 
                #full_width = F) %>% 
 #row_spec(39, bold = T)
```

### II. Document feature matrix
```{r}
# create document feature matrix
dfmat <- raw %>%
 select(text) %>% 
 unlist() %>% 
 tokens(remove_punc = TRUE) %>% 
 tokens_remove(pattern=stopwords("en")) %>% 
 tokens_replace(pattern = lexicon::hash_lemmas$token, replacement = lexicon::hash_lemmas$lemma) %>%
 dfm()

# rename dfm rows with meaningful IDs
rownames(dfmat) <- raw$doc_id

dfmat
```

```{r}
topfeatures(dfmat, n = 10, scheme = "docfreq")
```

### III. LDA 
```{r}
lda <- LDA(dfmat, control = list(alpha = 0.3, seed = 0), k = 12)
```

```{r}
# document-topic probability
topics <- tidy(lda, matrix = "gamma") %>% 
 left_join(., raw %>% select(!text), by = c("document" = "doc_id"))
```

##### By platform
```{r}
# topic by platform
share_company <- topics %>% 
 filter(year %in% c(2012:2022)) %>% 
 group_by(company_search, topic) %>% 
 summarise(gamma = sum(gamma)) %>% 
 group_by(company_search) %>% 
 mutate(share = gamma/sum(gamma)) %>% 
 ungroup() %>% 
 mutate(topic = factor(topic))
```

Document-topic probabilities: share by platform
```{r}
 plot1 <- ggplot(share_company, aes(x = topic, y = fct_rev(company_search), fill = share)) +
  geom_tile() +
  scale_fill_gradient2(high = "darkred", labels=scales::label_percent(accuracy = 1L)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank(),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 10, face = "italic", 
                                     hjust = 0.5, margin = margin(b = 15)),
        plot.caption = element_text(size = 10, hjust = 0),
        axis.title.y.left = element_text(size = 12, margin = margin(r = 15)),
        axis.title.x = element_text(size = 12, margin = margin(t = 15))) +
  labs(x = "Topics",
       y = "Platform",
       fill = "% share",
       title = "Share of Topics by Platform",
       subtitle = "LDA with 12 topics") 
plot1

ggsave(filename="lda_share_platforms.png", plot=plot1, device="png", 
       path=path_charts, width = 10, height = 7)

```

Uniqueness of terms
```{r}
# topic-word probability (words most common in each topic)
top_words <- tidy(lda, matrix = "beta") %>% 
 group_by(topic) %>% 
 slice_max(beta, n = 50) %>% 
 ungroup() %>%
 arrange(topic, -beta)
top_words

# categorize top words as unique or common 
words <- top_words %>% 
 group_by(term) %>% 
 summarise(counts = n()) %>% 
 arrange(counts) %>% 
 mutate(unique = ifelse(counts == 1, "unique", "common"))

words %>% 
 group_by(unique) %>% 
 summarise(counts = n())

# df with word classification
words_df <- left_join(top_words, words)
```

```{r}
plot2 <- ggplot(words_df %>% distinct(term, unique, .keep_all = TRUE), aes(x = unique, beta)) +
 geom_jitter(aes(color = unique), size = 4, alpha = 0.7) + 
 scale_color_manual(values = c("#ba0000", "#234075")) + 
 scale_alpha_manual(values = c(0.6, 0.6)) + 
 xlab("") + 
 ylab("beta") + 
 scale_y_continuous(limits = c(0, 0.03)) + 
 guides(alpha = "none") + 
 labs(title = "Common vs Unique Words",
      subtitle = "top 50 words per topic") + 
 guides(color = "none") + 
 charts.theme
plot2

ggsave(filename="lda_common_vs_unique.png", plot=plot2, device="png", 
       path=path_charts, width = 10, height = 7)
```
**More unique words but higher probability of common words occurring.**

Topic-word probabilities
```{r}
unique_words <- words_df %>% 
 mutate(non_an = grepl("[^\\p{L} ]", term, perl = TRUE)) %>% 
 filter(unique == "unique", non_an == FALSE) %>% 
 group_by(topic) %>% 
 slice_max(beta, n=10) %>% 
 ungroup() %>% 
 arrange(topic, beta)
```

```{r}
plot3 <-  ggplot(unique_words %>% filter(!(topic %in% c(2, 12))), aes(beta, reorder(term, beta))) +
  geom_col(show.legend = FALSE, fill = "#234075") +
  facet_wrap(~ topic, scales = "free", nrow = 2) +
  labs(title = "Top 10 words per topic") + 
  xlab("beta") + 
  ylab("") +
  charts.theme + 
  scale_y_reordered()
plot3

ggsave(filename="lda_words.png", plot=plot3, device="png", 
       path=path_charts, width = 18, height = 7)
```


##### By year
```{r}
# topic by year
share_period <- topics %>% 
 filter(year %in% c(2012:2022)) %>% 
 group_by(year, topic) %>% 
 summarise(gamma = sum(gamma)) %>% 
 group_by(year) %>% 
 mutate(share = gamma/sum(gamma)) %>% 
 ungroup() %>% 
 mutate(topic = factor(topic))
```

Document-topic probabilities: share by year
```{r}
 plot4 <- ggplot(share_period, aes(x = topic, y = as.factor(year), fill = share)) +
  geom_tile() +
  scale_fill_gradient2(high = "darkred", labels=scales::label_percent(accuracy = 1L)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank(),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 10, face = "italic", 
                                     hjust = 0.5, margin = margin(b = 15)),
        plot.caption = element_text(size = 10, hjust = 0),
        axis.title.y.left = element_text(size = 12, margin = margin(r = 15)),
        axis.title.x = element_text(size = 12, margin = margin(t = 15))) +
  labs(x = "Topics",
       y = "Year",
       fill = "% share",
       title = "Share of Topics by Year",
       subtitle = "LDA with 12 topics") 
plot4

ggsave(filename="lda_share_year.png", plot=plot4, device="png", 
       path=path_charts, width = 10, height = 7)
```

##### By region
**Note: smaller dataset**
```{r}
continents <- read_csv(paste0(path_repo, "/data/others/continents.csv")) %>% 
 select(name, region)
colnames(continents)[1] <- "country"

topics_region <- left_join(topics, continents) %>% 
 filter(!is.na(region)) %>% 
 mutate(eu = ifelse(region == "Europe", "Europe", "Outside Europe"))

topics_region %>% 
 group_by(eu) %>% 
 summarise(counts = n())
```

Document-topic probabilities: share by region
```{r}
share_eu <- topics_region %>% 
 group_by(eu, topic) %>% 
 summarise(gamma = sum(gamma)) %>% 
 group_by(eu) %>% 
 mutate(share = gamma/sum(gamma)) %>% 
 ungroup() %>% 
 mutate(topic = factor(topic))

region_df <- share_eu %>% 
 group_by(eu) %>% 
 mutate(alpha = ifelse(max(share) == share, "yes", "no"), 
        share = ifelse(eu == "Europe", -1*share, share))
```

```{r}
plot5 <-  ggplot(region_df %>% filter(abs(share) > 0.04), aes(x=fct_rev(topic), y=share*100, label=eu)) + 
  geom_hline(yintercept = 0, size = 0.4, linetype = 2) +
  geom_point(stat='identity', fill="black", size=6, aes(color = eu, alpha = alpha))  +
  geom_segment(aes(y = 0, 
                   x = topic, 
                   yend = ifelse(eu == "Europe", share*100 + 0.5, share*100 - 0.5), 
                   xend = topic, 
                   color = eu,
                   alpha = alpha),
               size = 1.5) +
  labs(title="Share of Topics by Region", 
       subtitle="Europe vs Outside Europe",
       caption="Note: Only for documents with mention of a country in the title or text. 
       Topics with less than 3% share removed.",
       color = "") + 
  xlab("topic") + 
  ylab("% share of topic") + 
  scale_y_continuous(labels = abs) + 
  scale_color_manual(values = c("#234075", "#ba0000")) + 
  scale_alpha_manual(values = c(0.6, 1)) + 
  guides(alpha = "none") + 
  charts.theme +
  coord_flip()
plot5

ggsave(filename="lda_share_eu.png", plot=plot5, device="png", 
       path=path_charts, width = 10, height = 7)
```

```{r}
share_region <- topics_region %>% 
 group_by(region, topic) %>% 
 summarise(gamma = sum(gamma)) %>% 
 group_by(region) %>% 
 mutate(share = gamma/sum(gamma)) %>% 
 ungroup() %>% 
 mutate(topic = factor(topic))
```

Document-topic probabilities: share by region
```{r}
 plot6 <- ggplot(share_region, aes(x = topic, y = fct_rev(region), fill = share)) +
  geom_tile() +
  scale_fill_gradient2(high = "darkred", labels=scales::label_percent(accuracy = 1L)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank(),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 10, face = "italic", 
                                     hjust = 0.5, margin = margin(b = 15)),
        plot.caption = element_text(size = 10, hjust = 0),
        axis.title.y.left = element_text(size = 12, margin = margin(r = 15)),
        axis.title.x = element_text(size = 12, margin = margin(t = 15))) +
  labs(x = "Topics",
       y = "Region",
       fill = "% share",
       title = "Share of Topics by Region",
       subtitle = "LDA with 12 topics") 
plot6

ggsave(filename="lda_share_region.png", plot=plot6, device="png", 
       path=path_charts, width = 10, height = 7)
```

