---
title: "Topic modeling: STM"
author: 
date: 
output:
  html_document:
    toc: yes
    keep_md: yes
    df_print: kable
    number_sections: no
    highlight: tango
    theme: lumen
    toc_depth: 3
    toc_float: yes
    css: custom.css
    self_contained: no
  pdf_document:
    toc: yes
    toc_depth: '3'
---
  
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>


```{r, include = F}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

<br>

***

```{r setup, include = T}
# loading packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(manifestoR, quanteda, tidyr, purrr, ggplot2, 
               tidytext, httr, rvest, readr, xml2, reshape2,
               stringr, stringi, dplyr, tibble, lexicon,
               NMF, topicmodels, LDAvis, stm)
```

```{r}
dir <- "D:/Desktop/02 Text as Data/antitRust"

df <- read.csv(file.path(dir, "data/df_clean_eng.csv")) %>%
  mutate(document = 1:4256)
```


```{r}
df %>% 
 group_by(year) %>%
 summarise(counts = n()) %>% 
 arrange(desc(year))

df %>% 
 filter(country != "None") %>% 
 group_by(country) %>%
 summarise(counts = n()) %>%
 arrange(desc(counts))

```

```{r}
processed <- textProcessor(documents=df$text,
                                 metadata = df,
                                 lowercase = TRUE, #*
                                 removestopwords = TRUE, #*
                                 removenumbers = TRUE, #*
                                 removepunctuation = TRUE, #*
                                 stem = TRUE, #*
                                 wordLengths = c(3,Inf), #*
                                 sparselevel = 1, #*
                                 language = "en", #*
                                 verbose = TRUE, #*
                                 onlycharacter = TRUE, # not def
                                 striphtml = FALSE, #*
                                 customstopwords = NULL, #*
                                 v1 = FALSE) #*

plotRemoved(processed$documents, lower.thresh=seq(1,200, by=100))

processed$vocab <- str_replace_all(processed$vocab, "[â|ä|ã|€|µ|º]", "")
out <- prepDocuments(processed$documents, processed$vocab, processed$meta, lower.thresh = 20)

```

# ```{r}
# #search no. of topics
# kResult <- searchK(out$documents, out$vocab, K=c(10,100), prevalence=~company_search+s(year),
#                    data=out$meta)
# ```
# 
# ```{r}
# png(file.path(dir, "/plots/stm_searchK.png"), width = 1000, height = 800)
# plot(kResult)
# dev.off()
# ```


```{r}
gafamfit <- stm(out$documents, out$vocab, K=20, prevalence=~company_search+s(year), 
                       max.em.its=75, data=out$meta, init.type="Spectral", 
                       seed=28)
```

```{r}
labelTopics(gafamfit)
```

```{r}
#par(bty="n",col="#de2d26",lwd=3)
png(file.path(dir, "/plots/stm_top_topics.png"), width = 1000, height = 800)
plot.STM(gafamfit,type="summary", n = 10 ,text.cex=1.2, width=80, height=90, xlim=c(0,.3)) #cex.main =1
dev.off()

png(file.path(dir, "/plots/stm_top_topics_unique.png"), width = 1000, height = 800)
plot.STM(gafamfit,type="summary", label="frex", n = 10 ,text.cex=1.2, width=80, height=90, xlim=c(0,.3)) #cex.main =1
dev.off()

plot.STM(gafamfit,type="summary", n = 10 ,text.cex=1.2, width=80, height=90, xlim=c(0,.3))
plot.STM(gafamfit,type="summary", label="frex", n = 10 ,text.cex=1.2, width=80, height=90, xlim=c(0,.3))
plot(gafamfit, type="labels", topics=c(4,8,14,17)) #label = "frex",
plot(gafamfit, type="labels", topics=c(6,16,3,12))
plot(gafamfit, type="labels", topics=c(11,15,19,20))
```

Topics (**interesting results)
Topic 4 - EC investigation on Reuter
Topic 8 - US FTC investigation on big tech industries brought to senate
Topic 14 - new business investment from Uber to Tiktok
Topic 17 - EC rules on fines
Topic 6 - 
**Topic 16 - Social media platforms for news content Facebook, Instagram, Meta, Twitter
Topic 3 - mergers and acquisitions in general
**Topic 12 - competition in innovative markets and platforms
**Topic 11 - online retail market in India (Amazon, Flipkart)
Topic 15 - Google search engine dominant position
Topic 19
**Topic 20 - app/game store payment system (Apple vs Fortnite Epic games)
**Topic 10 - music streaming and apple's mobile payment service 
**Topic 5 - data privacy and protection (Facebook, Whatsapp)
Topic 7 - competition enforcement in general
Topic 1 - mobile device/smartphone's operating system 
Topic 13 - digital markets act
Topic 2 - Microsoft cloud services
Topic 18 - online advertising Google, Youtube
**Topic 9 - patent and licensing mobile phones (Qualcomm, Samsung, Nokia, Motorola)

## Topic Correlation

```{r}
# correlation of the topics
corr <- topicCorr(gafamfit)

# png(file.path(dir, "/plots/stm_topic_correlation.png"), width = 1000, height = 800)
# plot(corr)
# dev.off()

plot(corr)
plot(gafamfit, type="labels", topics=c(1,8,7))
plot(gafamfit, type="labels", topics=c(3,6,5))
```

## Doc-topic charts
```{r}
doc_topic <- tidy(gafamfit, matrix = "theta") %>%
  left_join(out$meta, by="document")

# topic by platform
share_company <- doc_topic %>% 
 filter(year %in% c(2012:2022)) %>% 
 group_by(company_search, topic) %>% 
 summarise(gamma = sum(gamma)) %>% 
 group_by(company_search) %>% 
 mutate(share = gamma/sum(gamma)) %>% 
 ungroup() %>% 
 mutate(topic = factor(topic))

plot1 <- ggplot(share_company, aes(x = topic, y = company_search, fill = share)) +
  geom_tile() +
  scale_fill_gradient2(high = "darkred", labels=scales::label_percent(accuracy = 1L)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank(),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 10, face = "italic", 
                                     hjust = 0.5, margin = margin(b = 15)),
        plot.caption = element_text(size = 10, hjust = 0),
        axis.title.y.left = element_text(size = 12, margin = margin(r = 15)),
        axis.title.x = element_text(size = 12, margin = margin(t = 15))) +
  labs(x = "Topics",
       y = "Platform",
       fill = "% share",
       title = "Share of Topics by Platform",
       subtitle = "STM with 20 topics") 
plot1
# ggsave(filename=file.path(dir, "plots/stm_share_platforms.png"), plot=plot1, device="png", 
#        width = 12, height = 7)

```


```{r}
# topic by year
share_period <- doc_topic %>%
  filter(year %in% c(2012:2022)) %>% 
  group_by(year, topic) %>%
  summarise(gamma = sum(gamma)) %>%
  group_by(year) %>%
  mutate(year_share = gamma/sum(gamma)) %>%
  ungroup() %>%
  mutate(topic = factor(topic))

plot2 <- ggplot(share_period, aes(x = topic, y = as.factor(year), fill = year_share)) +
  geom_tile() +
  scale_fill_gradient2(high = "darkred", labels=scales::label_percent(accuracy = 1L)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank(),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 10, face = "italic", 
                                     hjust = 0.5, margin = margin(b = 15)),
        plot.caption = element_text(size = 10, hjust = 0),
        axis.title.y.left = element_text(size = 12, margin = margin(r = 15)),
        axis.title.x = element_text(size = 12, margin = margin(t = 15))) +
  labs(x = "Topics",
       y = "Year",
       fill = "% share",
       title = "Share of Topics by Year",
       subtitle = "STM with 20 topics") 
plot2
# ggsave(filename= file.path(dir, "plots/stm_share_year.png"), plot=plot2, device="png", 
#        width = 12, height = 7)

```

## Topic-term charts
```{r}
topic_term <- tidy(gafamfit, matrix = "beta")

term_share <- topic_term %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>% 
  ungroup() %>%
  arrange(topic, -beta)

plot3 <-term_share %>%
      mutate(term = reorder_within(term, beta, topic)) %>%
      ggplot(aes(beta, term, fill = "red")) +
      geom_bar(stat="identity", width = .5) +
      facet_wrap(~ topic, scales = "free", ncol=5) + 
      #xlim(0, 0.08) +
      scale_y_reordered() +
      labs(title = "STM") +
      scale_fill_identity() +
      theme_test() +
      theme(axis.text.y = element_text(size = 8),
            axis.text.x = element_text(size = 8),
            strip.text.x = element_text(size = 8)) +
    labs(y = "Term", x = "Term probability per topic") 

plot3

ggsave(filename= file.path(dir, "plots/stm_term_share.png"), plot=plot3, device="png",
       width = 12, height = 7)
```


```{r}
df_fulltext <- df$text
df_shorttext <- substr(df$text,1,500)

findThoughts(gafamfit, texts=df_shorttext, n=3, topics=20)

#df %>% filter(str_detect(title, "US: Microsoft, Oracle gear up for latest non-poaching fight"))

```

## Estimate Effect

```{r}
out$meta$company_search <- as.factor(out$meta$company_search)
prep <- estimateEffect(1:20 ~ company_search+s(year), gafamfit, meta=out$meta, 
                       uncertainty="Global")

summary(prep, topics=c(20))
```


```{r}
plot(prep, covariate="company_search", topics=c(20), model=gafamfit, 
     method="difference", cov.value1="Amazon", cov.value2="Microsoft",
     xlab="", main="Effect of Company",
     xlim=c(-.15,.15)) #labeltype ="custom", custom.labels=c())

```


```{r}
plot(gafamfit, type="perspectives", topics=c(7,10))
     
cloud(gafamfit, topic=7)

```




```{r}
topicQuality(model=gafamfit, documents=out$documents)
```



```{r}
# english_words <- readLines("D:/Documents/R/win-library/4.1/hunspell/dict/en_US.dic") %>% 
#   gsub("/.+", "", .)
# 
# dfmat <- df$text %>%
#         tokens(remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE) %>%
#         tokens_remove(pattern=stopwords("en")) %>%
#         tokens_remove("[^[:alnum:]]") %>%
#         tokens_keep(english_words, valuetype = "fixed") %>% 
#         #tokens_remove(omit_words) %>%
#         tokens_wordstem() %>%
#         dfm()  %>%
#         dfm_trim(min_termfreq = 10) 
# 
# dfm_stm <- convert(dfmat, to = "stm")

# plotRemoved(dfm_stm$documents, lower.thresh=seq(1,200, by=100))
# 
# 
# model <- stm(documents = dfm_stm$documents,
#            vocab = dfm_stm$vocab,
#            K = 10,
#            verbose = TRUE)

```

```{r}
# labelTopics(model)
# 
# 
# par(bty="n",col="#de2d26",lwd=3)
# plot.STM(model,type="summary", label="frex", n = 3, main = "2020 Republicans", text.cex=.8, width=30, cex.main =.8)

```
